---
title: "Fecal transplantation piglet project"
author: "Ming Yan"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```



# Load the libraries 

```{r}
library(tidyverse)
library(stringr)
library(vegan)
library(phyloseq)
library(ape)
library(glue)
library(ggpubr)
library(rstatix)
library(patchwork)
library(ANCOMBC)
library(ComplexHeatmap)
```

# Data import and formating

## Import QIIME2 output and create phyloseq objectives

```{r}
otu <- read.table(file = "feature-table.tsv", sep = "\t", header = T, row.names = 1, 
                   comment.char = "")
taxonomy <- read.table(file = "taxonomy.tsv", sep = "\t", header = T ,row.names = 1)
tax <- taxonomy %>%
  select(Taxon) %>% 
  separate(Taxon, c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species"), "; ")


tax.clean <- data.frame(row.names = row.names(tax),
                        Kingdom = str_replace(tax[,1], "k__",""),
                        Phylum = str_replace(tax[,2], "p__",""),
                        Class = str_replace(tax[,3], "c__",""),
                        Order = str_replace(tax[,4], "o__",""),
                        Family = str_replace(tax[,5], "f__",""),
                        Genus = str_replace(tax[,6], "g__",""),
                        Species = str_replace(tax[,7], "s__",""),
                        stringsAsFactors = FALSE)

tax.clean[is.na(tax.clean)] <- ""
tax.clean[tax.clean=="__"] <- ""

for (i in 1:nrow(tax.clean)){
  if (tax.clean[i,7] != ""){
    tax.clean$Species[i] <- paste(tax.clean$Genus[i], tax.clean$Species[i], sep = " ")
  } else if (tax.clean[i,2] == ""){
    kingdom <- paste("Unclassified", tax.clean[i,1], sep = " ")
    tax.clean[i, 2:7] <- kingdom
  } else if (tax.clean[i,3] == ""){
    phylum <- paste("Unclassified", tax.clean[i,2], sep = " ")
    tax.clean[i, 3:7] <- phylum
  } else if (tax.clean[i,4] == ""){
    class <- paste("Unclassified", tax.clean[i,3], sep = " ")
    tax.clean[i, 4:7] <- class
  } else if (tax.clean[i,5] == ""){
    order <- paste("Unclassified", tax.clean[i,4], sep = " ")
    tax.clean[i, 5:7] <- order
  } else if (tax.clean[i,6] == ""){
    family <- paste("Unclassified", tax.clean[i,5], sep = " ")
    tax.clean[i, 6:7] <- family
  } else if (tax.clean[i,7] == ""){
    tax.clean$Species[i] <- paste("Unclassified ",tax.clean$Genus[i], sep = " ")
  }
}


metadata <- read.table(file = "metadata.tsv", sep = "\t", header = T, row.names = 1)
metadata$'sample.id' <- rownames(metadata)

##rename
metadata$time <- str_replace(metadata$time, "Pre_inoculation" , "Day0 (Pre-inoculation)")
metadata$time <- str_replace(metadata$time, "DPC0" , "Day40 (DPC0)")
metadata$time <- str_replace(metadata$time, "DPC2" , "Day42 (DPC2)")
metadata$time <- str_replace(metadata$time, "DPC4" , "Day44 (DPC4)")
metadata$time <- str_replace(metadata$time, "DPC7" , "Day47 (DPC7)")

metadata_con <- metadata %>% filter(Treatment == "Control")
metadata<- metadata %>% filter(Experiment.id == 2) ## only use data from experiment 2, because immunology data from experiment 2
colnames(otu) <- str_replace_all(colnames(otu),"X","")
```

## Build phyloseq objectives
```{r}
OTU = otu_table(as.matrix(otu), taxa_are_rows = TRUE)
TAX = tax_table(as.matrix(tax.clean))
SAMPLE <- sample_data(metadata)
SAMPLE_CON <- sample_data(metadata_con)
ps <- phyloseq(OTU, TAX, SAMPLE)
ps_control <- phyloseq(OTU, TAX, SAMPLE_CON)
```

## Rarefaction
```{r}
set.seed(111) # keep result reproductive
ps.rarefied = rarefy_even_depth(ps, rngseed=1, sample.size=4000, replace=F)
ps.rarefied.con = rarefy_even_depth(ps_control, rngseed=1, sample.size=4000, replace=F)

```

## Format metadata
```{r}
meta_stats <- metadata %>% filter(type != "Buffer")

meta_stats <- meta_stats %>% tibble::rownames_to_column("sample-id")
```

# Alpha diversity

## calculate diversity index
```{r}
alpha = estimate_richness(ps.rarefied, measures = c("Observed", "Shannon","Chao1","Shannon","Simpson"))
rownames(alpha) <- str_replace_all(rownames(alpha),"X","")
alpha1 <- alpha %>% tibble::rownames_to_column("sample-id")
alpha_ready <- merge(alpha1,meta_stats, by = "sample-id")
alpha_fecal <- alpha_ready %>% filter(type == "Fecal") %>% filter(Treatment != 'Control')
```

## Plot shannon
```{r}
# Shannon
shannon <- alpha_fecal %>% ggplot(aes(x=Treatment, y=Shannon)) + 
  geom_boxplot() +
  facet_wrap(~time) +
  geom_jitter()

shannon <- shannon + theme_classic() +  ## remove panel.border
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        panel.border = element_rect(colour = "black", fill = NA)) +
   labs(x = "") +
   labs(y = "Shannon")

shannon
#ggsave('res_ex2/shannon.png', plot = shannon,device = 'png',width = 6, height = 4)
```


## Stats for alpha
```{r}

# shannon <- alpha_fecal %>% select("Treatment","Shannon","time") %>%  group_by(time) %>% do(w = wilcox.test(data=., Shannon~ Treatment, paired=FALSE)) %>% summarise(time, Wilcox = w$p.value) 

shannon <- alpha_fecal %>% select("Treatment","Shannon","time") %>%  group_by(time) %>% wilcox_test(formula = Shannon~ Treatment, paired=FALSE) %>%
  add_significance()  

simpson <- alpha_fecal %>% select("Treatment","Simpson","time") %>%  group_by(time) %>% wilcox_test(formula = Simpson~ Treatment, paired=FALSE) %>%
  add_significance()  

richness <- alpha_fecal %>% select("Treatment","Observed","time") %>%  group_by(time) %>% wilcox_test(formula = Observed~ Treatment, paired=FALSE) %>%
  add_significance() 



shannon
simpson
richness

```
*none is different*

# Beta diversity

## PCoA 
```{r}
control <- ps.rarefied.con %>% subset_samples(type == 'Fecal')
ps_fecal_ex2 <- ps.rarefied %>% subset_samples(type == 'Fecal')
ps_fecal_ex2 <- merge_phyloseq(ps_fecal_ex2,control)


bray_ex2 <- phyloseq::distance(ps_fecal_ex2, method="bray")

pcoa_ex2 <- ordinate(ps_fecal_ex2, method="PCoA", distance=bray_ex2)
ex2 <- plot_ordination(ps_fecal_ex2 , pcoa_ex2, color="Treatment")  + 
  theme_classic() +
  theme(strip.background = element_blank()) +
  facet_wrap(~time, scales="free")
leg_ex2 <- get_legend(ex2)  # from library "ggpubr"
  
fecal_pre_ex2 <- ps_fecal_ex2 %>% subset_samples(time == 'Day0 (Pre-inoculation)')  ## before virus challenge
bray_pre_ex2 <- phyloseq::distance(fecal_pre_ex2, method="bray") 
pcoa_pre_ex2 <- ordinate(fecal_pre_ex2, method="PCoA", distance=bray_pre_ex2)
pre_ex2 <- plot_ordination(fecal_pre_ex2, pcoa_pre_ex2, color="Treatment") + 
  theme_classic()  + theme(plot.title = element_text(hjust = 0.5)) +labs(title = "Pre-inoculation", x = paste("PC1 [", round(pcoa_pre_ex2$values$Relative_eig[1]*100, 1), "%]",sep = ""),
    y = paste("PC2 [", round(pcoa_pre_ex2$values$Relative_eig[2]*100, 1), "%]",sep = "")) + theme(legend.position = "none")


fecal_22_ex2 <- ps_fecal_ex2 %>% subset_samples(time == 'Day22')  ## before virus challenge
bray_22_ex2 <- phyloseq::distance(fecal_22_ex2, method="bray") 
pcoa_22_ex2 <- ordinate(fecal_22_ex2, method="PCoA", distance=bray_22_ex2)
day_22_ex2 <- plot_ordination(fecal_22_ex2, pcoa_22_ex2, color="Treatment") + 
  theme_classic()  + theme(legend.position = "none") +labs(title = "Day 22", x = paste("PC1 [", round(pcoa_22_ex2$values$Relative_eig[1]*100, 1), "%]",sep = ""),
    y = paste("PC2 [", round(pcoa_22_ex2$values$Relative_eig[2]*100, 1), "%]",sep = "")) + 
  theme(plot.title = element_text(hjust = 0.5))


fecal_DPC0_ex2 <- ps_fecal_ex2 %>% subset_samples(time == 'Day40 (DPC0)')  ## before virus challenge
bray_DPC0_ex2 <- phyloseq::distance(fecal_DPC0_ex2, method="bray") 
pcoa_DPC0_ex2 <- ordinate(fecal_DPC0_ex2, method="PCoA", distance=bray_DPC0_ex2)
DPC0_ex2 <-plot_ordination(fecal_DPC0_ex2, pcoa_DPC0_ex2, color="Treatment") + 
  theme_classic()  + theme(plot.title = element_text(hjust = 0.5),
    legend.position = "none") +labs(title = "DPC0 (day 40)", x = paste("PC1 [", round(pcoa_22_ex2$values$Relative_eig[1]*100, 1), "%]",sep = ""),
    y = paste("PC2 [", round(pcoa_22_ex2$values$Relative_eig[2]*100, 1), "%]",sep = ""))

fecal_DPC2_ex2 <- ps_fecal_ex2 %>% subset_samples(time == 'Day42 (DPC2)')  ## before virus challenge
bray_DPC2_ex2 <- phyloseq::distance(fecal_DPC2_ex2, method="bray") 
pcoa_DPC2_ex2 <- ordinate(fecal_DPC2_ex2, method="PCoA", distance=bray_DPC2_ex2)
DPC2_ex2 <- plot_ordination(fecal_DPC2_ex2, pcoa_DPC2_ex2, color="Treatment") + 
  theme_classic()  + theme(plot.title = element_text(hjust = 0.5),
    legend.position = "none") +labs(title = "DPC2 (day 42)", x = paste("PC1 [", round(pcoa_DPC2_ex2$values$Relative_eig[1]*100, 1), "%]",sep = ""),
    y = paste("PC2 [", round(pcoa_DPC2_ex2$values$Relative_eig[2]*100, 1), "%]",sep = ""))

fecal_DPC4_ex2 <- ps_fecal_ex2 %>% subset_samples(time == 'Day44 (DPC4)')  ## before virus challenge
bray_DPC4_ex2 <- phyloseq::distance(fecal_DPC4_ex2, method="bray") 
pcoa_DPC4_ex2 <- ordinate(fecal_DPC4_ex2, method="PCoA", distance=bray_DPC4_ex2)
DPC4_ex2 <- plot_ordination(fecal_DPC4_ex2, pcoa_DPC4_ex2, color="Treatment") + 
  theme_classic()  + theme(plot.title = element_text(hjust = 0.5),
    legend.position = "none") +labs(title = "DPC4 (day 44)", x = paste("PC1 [", round(pcoa_DPC4_ex2$values$Relative_eig[1]*100, 1), "%]",sep = ""),
    y = paste("PC2 [", round(pcoa_DPC4_ex2$values$Relative_eig[2]*100, 1), "%]",sep = ""))

fecal_DPC7_ex2 <- ps_fecal_ex2 %>% subset_samples(time == 'Day47 (DPC7)')  ## before virus challenge
bray_DPC7_ex2 <- phyloseq::distance(fecal_DPC7_ex2, method="bray") 
pcoa_DPC7_ex2 <- ordinate(fecal_DPC7_ex2, method="PCoA", distance=bray_DPC7_ex2)
DPC7_ex2 <- plot_ordination(fecal_DPC7_ex2, pcoa_DPC7_ex2, color="Treatment") + 
  theme_classic()  + theme(plot.title = element_text(hjust = 0.5),
    legend.position = "none") +labs(title = "DPC7 (day 47)", x = paste("PC1 [", round(pcoa_DPC7_ex2$values$Relative_eig[1]*100, 1), "%]",sep = ""),
    y = paste("PC2 [", round(pcoa_DPC7_ex2$values$Relative_eig[2]*100, 1), "%]",sep = ""))

ggarrange(ggarrange(pre_ex2,day_22_ex2, DPC0_ex2,DPC2_ex2,DPC4_ex2,DPC7_ex2),leg_ex2, widths = c(2.5, 0.5),
          ncol = 2, nrow = 1) 

#ggsave('PCoa_ex2.png', plot = ex2_by_time,device = 'png',width = 12, height = 4)
```


## Add Adonis results as the annotation
```{r}
adonis_pre_ex2 <- adonis(phyloseq::distance(fecal_pre_ex2 %>% subset_samples(Treatment %in% c("Rural", "Urban")), method="bray")  ~ Treatment, fecal_pre_ex2 %>% subset_samples(Treatment %in% c("Rural", "Urban")) %>% sample_data() %>% data.frame())

adonis_22_ex2 <- adonis(phyloseq::distance(fecal_22_ex2 %>% subset_samples(Treatment %in% c("Rural", "Urban")), method="bray")  ~ Treatment, fecal_22_ex2 %>% subset_samples(Treatment %in% c("Rural", "Urban")) %>% sample_data() %>% data.frame())

adonis_DPC0_ex2 <- adonis(phyloseq::distance(fecal_DPC0_ex2 %>% subset_samples(Treatment %in% c("Rural", "Urban")), method="bray")  ~ Treatment, fecal_DPC0_ex2 %>% subset_samples(Treatment %in% c("Rural", "Urban")) %>% sample_data() %>% data.frame())

adonis_DPC2_ex2 <- adonis(phyloseq::distance(fecal_DPC2_ex2 %>% subset_samples(Treatment %in% c("Rural", "Urban")), method="bray")  ~ Treatment, fecal_DPC2_ex2 %>% subset_samples(Treatment %in% c("Rural", "Urban")) %>% sample_data() %>% data.frame())

adonis_DPC4_ex2 <- adonis(phyloseq::distance(fecal_DPC4_ex2 %>% subset_samples(Treatment %in% c("Rural", "Urban")), method="bray")  ~ Treatment, fecal_DPC4_ex2 %>% subset_samples(Treatment %in% c("Rural", "Urban")) %>% sample_data() %>% data.frame())

adonis_DPC7_ex2 <- adonis(phyloseq::distance(fecal_DPC7_ex2 %>% subset_samples(Treatment %in% c("Rural", "Urban")), method="bray")  ~ Treatment, fecal_DPC7_ex2 %>% subset_samples(Treatment %in% c("Rural", "Urban")) %>% sample_data() %>% data.frame())


pre_ex2_anno <- ggplot() + 
  annotate("text", x = 2, y = 5, label = paste("Rural vs. Urban: ", adonis_pre_ex2$aov.tab$`Pr(>F)`[1], sep = "") , size = 3) +
  coord_cartesian(clip = "off") +   # allows text to overflow if needed
  theme_void()

pre_ex2_final <- pre_ex2 + inset_element(pre_ex2_anno, left = 0.55, bottom = 1.05, right = 0.55, top = 1) # function "inset_element" from library "patchwork" 


D22_ex2_anno <- ggplot() + 
  annotate("text", x = 2, y = 5, label = paste("Rural vs. Urban: ", adonis_22_ex2$aov.tab$`Pr(>F)`[1], sep = "") , size = 3) +
  coord_cartesian(clip = "off") +   # allows text to overflow if needed
  theme_void()

D22_ex2_final <- day_22_ex2 + inset_element(D22_ex2_anno, left = 0.55, bottom = 1.05, right = 0.55, top = 1)


DPC0_ex2_anno <- ggplot() + 
  annotate("text", x = 2, y = 5, label = paste("Rural vs. Urban: ", adonis_DPC0_ex2$aov.tab$`Pr(>F)`[1], sep = "") , size = 3) +
  coord_cartesian(clip = "off") +   # allows text to overflow if needed
  theme_void()

DPC0_ex2_final <- DPC0_ex2 + inset_element(DPC0_ex2_anno, left = 0.55, bottom = 1.05, right = 0.55, top = 1)


DPC2_ex2_anno <- ggplot() + 
  annotate("text", x = 2, y = 5, label = paste("Rural vs. Urban: ", adonis_DPC2_ex2$aov.tab$`Pr(>F)`[1], sep = "") , size = 3) +
  coord_cartesian(clip = "off") +   # allows text to overflow if needed
  theme_void()

DPC2_ex2_final <- DPC2_ex2 + inset_element(DPC2_ex2_anno, left = 0.55, bottom = 1.05, right = 0.55, top = 1)


DPC4_ex2_anno <- ggplot() + 
  annotate("text", x = 2, y = 5, label = paste("Rural vs. Urban: ", adonis_DPC4_ex2$aov.tab$`Pr(>F)`[1], sep = "") , size = 3) +
  coord_cartesian(clip = "off") +   # allows text to overflow if needed
  theme_void()

DPC4_ex2_final <- DPC4_ex2 + inset_element(DPC4_ex2_anno, left = 0.55, bottom = 1.05, right = 0.55, top = 1)


DPC7_ex2_anno <- ggplot() + 
  annotate("text", x = 2, y = 5, label = paste("Rural vs. Urban: ", adonis_DPC7_ex2$aov.tab$`Pr(>F)`[1], sep = "") , size = 3) +
  coord_cartesian(clip = "off") +   # allows text to overflow if needed
  theme_void()

DPC7_ex2_final <- DPC7_ex2 + inset_element(DPC7_ex2_anno, left = 0.55, bottom = 1.05, right = 0.55, top = 1)




pcoa_ex2 <-  ggarrange(ggarrange(pre_ex2_final, D22_ex2_final, DPC0_ex2_final, DPC2_ex2_final, DPC4_ex2_final, DPC7_ex2_final),leg_ex2, widths = c(2.5, 0.5),
          ncol = 2, nrow = 1) 

pcoa_ex2
ggsave('res_ex2/pcoa.png', plot = pcoa_ex2,device = 'png',width = 8, height = 6)

```

# Phylum level Abundance bar plot

## Fecal
```{r}
rela_fecal_ex2 <- transform_sample_counts(ps_fecal_ex2, function(x) x / sum(x) )
glom_ex2 <- tax_glom(rela_fecal_ex2, taxrank = 'Phylum', NArm = FALSE)
ps.melt_ex2 <- psmelt(glom_ex2)
# change to character for easy-adjusted level
ps.melt_ex2$Phylum <- as.character(ps.melt_ex2$Phylum)

ps.melt_ex2 <- ps.melt_ex2 %>%
  group_by(time, Phylum) %>%
  mutate(median=median(Abundance))
# select group median > 1%
keep_ex2 <- unique(ps.melt_ex2$Phylum[ps.melt_ex2$median > 0.01])
ps.melt_ex2$Phylum[!(ps.melt_ex2$Phylum %in% keep_ex2)] <- "Other phyla (< 1%)"

rela_ex2_fecal <- ps.melt_ex2 %>% group_by(Treatment,time, Phylum) %>% summarise(Abundance=mean(Abundance)) 

rela_ex2_fecal$time <- as.factor(rela_ex2_fecal$time)
levels(rela_ex2_fecal$time) <- c("Day0 (Pre-inoculation)", "Day22", "Day40 (DPC0)","Day42 (DPC2)","Day44 (DPC4)","Day47 (DPC7)") # reorder

## create color palette for barplots
mypalette <- c("#FF0000", "#FFDB00", "#1bbf06", "#42f5d7", "#0092FF", "#4900FF", "#FF00DB")
names(mypalette) <- unique(rela_ex2_fecal$Phylum)


barplot_ex2_fecal <- ggplot(rela_ex2_fecal, aes(x = Treatment, y = Abundance, fill = Phylum)) + 
geom_bar(stat = "identity", aes(fill=Phylum)) + 
labs(x="", y="Fecal microbiota") +
theme_classic() + 
  scale_fill_manual(values = mypalette) +
theme(strip.background = element_blank(), 
        axis.text.x.bottom = element_text(angle = 90)) +
  facet_wrap(~time) +
  theme(legend.position = "none") +
  theme(legend.direction="horizontal")

legend_abundance <- get_legend( ggplot(rela_ex2_fecal, aes(x = Treatment, y = Abundance, fill = Phylum)) + 
geom_bar(stat = "identity", aes(fill=Phylum)) + 
labs(x="", y="Fecal microbiota") +
theme_classic() + 
  scale_fill_manual(values = mypalette) +
  theme(legend.direction="horizontal"))  

barplot_ex2_fecal
#ggsave('barplot_fecal_ex2.png', plot = barplot_ex2_fecal,device = 'png',width = 8, height = 4)
```


## Digestive tract 
```{r}
## control filtered after rarefaction
ps_dig <- ps.rarefied %>% subset_samples(type %in% c("Ileum", "Colon"))  
rela_dig <- transform_sample_counts(ps_dig, function(x) x / sum(x) ) 

dig_glom <- tax_glom(rela_dig, taxrank = 'Phylum', NArm = FALSE)
dig.melt <- psmelt(dig_glom)
# change to character for easy-adjusted level
dig.melt$Phylum <- as.character(dig.melt$Phylum)

dig.melt <- dig.melt %>%
  group_by(time, type,  Phylum) %>%
  mutate(median=median(Abundance))
# select group median > 1%
keep_dig <- unique(dig.melt$Phylum[dig.melt$median > 0.01])
dig.melt$Phylum[!(dig.melt$Phylum %in% keep_dig)] <- "Low abundance (< 1%)"

rela_dig_ready <- dig.melt %>% group_by(Treatment,time, type, Phylum) %>% summarise(Abundance=mean(Abundance)) 

rela_dig_ready$time <- as.factor(rela_dig_ready$time)


barplot_dig_by_time <- ggplot(rela_dig_ready, aes(x = Treatment, y = Abundance, fill = Phylum)) + 
geom_bar(stat = "identity", aes(fill=Phylum)) + 
labs(x="", y="Intestinal microbiota") +
theme_classic() + 
  scale_fill_manual(values = mypalette) +
theme(strip.background = element_blank(), 
        axis.text.x.bottom = element_text(angle = 90)) +
  facet_wrap(~time + type) +
  theme(legend.position = "none")
barplot_dig_by_time



#ggsave('barplot_nasal.png', plot = barplot_nasal_by_trt,device = 'png',width = 8, height = 4)
```


## Respiration tract
```{r}
ps_res <- ps.rarefied %>% subset_samples(type %in% c('Nasal', "BAL"))  
rela_res <- transform_sample_counts(ps_res, function(x) x / sum(x) ) 
#nasal_filter <- filter_taxa(physeq = rela_nasal, function(x) sum(x>1e-4) >(0.10*length(x)) , TRUE) %>% transform_sample_counts( function(x) x / sum(x) ) 

res_glom <- tax_glom(rela_res, taxrank = 'Phylum', NArm = FALSE)
res.melt <- psmelt(res_glom)
# change to character for easy-adjusted level
res.melt$Phylum <- as.character(res.melt$Phylum)

res.melt <- res.melt %>%
  group_by(time, Phylum) %>%
  mutate(median=median(Abundance))
# select group median > 1%
keep_res <- unique(res.melt$Phylum[res.melt$median > 0.01])
res.melt$Phylum[!(res.melt$Phylum %in% keep_res)] <- "Low abundance (< 1%)"

rela_res_ready <- res.melt %>% group_by(Treatment,time,type, Phylum) %>% summarise(Abundance=mean(Abundance)) 

rela_res_ready$time <- as.factor(rela_res_ready$time)
levels(rela_res_ready$time) <- c("Day0 (Pre-inoculation)","Day40 (DPC0)","Day42 (DPC2)","Day44 (DPC4)","Day47 (DPC7)", "Day22") # reorder


barplot_res_by_time <- ggplot(rela_res_ready, aes(x = Treatment, y = Abundance, fill = Phylum)) + 
geom_bar(stat = "identity", aes(fill=Phylum)) + 
labs(x="", y="Respiratory tissues microbiota") +
theme_classic() + 
  scale_fill_manual(values = mypalette) +
theme(strip.background = element_blank(), 
        axis.text.x.bottom = element_text(angle = 90)) +
  facet_wrap(~time + type ) +
  theme(legend.position = "none")
barplot_res_by_time 




#ggsave('barplot_nasal.png', plot = barplot_nasal_by_trt,device = 'png',width = 8, height = 4)
```



## Arrange as single plot
```{r}
phylum_abundance <- ggarrange(ggarrange(barplot_ex2_fecal, barplot_dig_by_time, barplot_res_by_time,  ncol = 1, nrow = 3, heights = c(1,1,1)),legend_abundance, heights = c(10, 1),
          ncol = 1, nrow = 2)

phylum_abundance
#ggsave('res_ex2/phylum_abundance.png', plot = phylum_abundance,device = 'png',width = 6, height = 12)
```


# Differential abundance analysis

##ANCOM BC

```{r}
`%!in%` <- Negate(`%in%`)
## preprocessing
 fecal_rela_filtered <- filter_taxa(physeq = rela_fecal_ex2, function(x) sum(x>1e-4) > (0.10*length(x)) , TRUE) %>% transform_sample_counts( function(x) x / sum(x) ) 
 fecal_taxa_filtered <- fecal_rela_filtered@otu_table %>% rownames()
fecal_ps_filtered <- prune_taxa(fecal_taxa_filtered, ps_fecal_ex2)
ancom_fecal_input <- tax_glom(fecal_ps_filtered %>% subset_samples(Treatment != 'Control') %>% subset_samples(time %!in% c('Pre_inoculation','Day22'))  , taxrank = 'Genus', NArm = FALSE)

## ANCOMBC
ancom_fecal_output <- ancombc(phyloseq = ancom_fecal_input, formula = "Treatment" , 
              p_adj_method = "holm", zero_cut = 0.70, lib_cut = 0, 
              group = "Treatment", struc_zero = TRUE, neg_lb = TRUE, tol = 1e-5, 
              max_iter = 100, conserve = TRUE, alpha = 0.05, global = FALSE)

sig_taxa <- ancom_fecal_output$res$diff_abn %>% as.data.frame() %>% filter(TreatmentUrban == "TRUE") %>% rownames()

## sig taxa rela abundance
fecal_genus_rela <- tax_glom(ps_fecal_ex2 %>% transform_sample_counts( function(x) x / sum(x))   , taxrank = 'Genus', NArm = FALSE)
fecal_genus_rela_sig <- prune_taxa(sig_taxa, fecal_genus_rela)
##

fecal_genus_table_sig <- t(fecal_genus_rela_sig@otu_table %>% as.data.frame())
sig_taxa_name <- fecal_genus_rela_sig@tax_table %>% as.data.frame() %>% select("Genus") %>% unlist() %>% as.vector()
colnames(fecal_genus_table_sig) <- sig_taxa_name
fecal_genus_metadata_sig <- fecal_genus_rela_sig@sam_data %>% as.data.frame()
fecal_genus_sig <- merge(fecal_genus_table_sig, fecal_genus_metadata_sig, by = "row.names") %>% pivot_longer(sig_taxa_name, names_to = "taxa", values_to = "abundance")%>% filter(Treatment != "Control")
fecal_genus_sig$abundance <- fecal_genus_sig$abundance*100
 
fecal_genus_sig$taxa <- str_replace_all(fecal_genus_sig$taxa, "Unclassified \\[Mogibacteriaceae\\]", "Unclassified Mogibacteriaceae") # formating
# plot
ancom_res <- ggplot(fecal_genus_sig, aes(x=Treatment, y=abundance,fill=Treatment)) +
  geom_boxplot()+ 
   facet_wrap(~taxa,scales= "free") + 
   labs(x = "", y = "Abundance (%)",
    fill = "Treatment")+  
   theme_classic() +  
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        panel.border = element_rect(colour = "black", fill = NA)) +
  #theme(legend.direction="horizontal")  +
  theme(legend.position = c(0.85, 0.2))

ancom_res
ggsave('res_ex2/ancom.png', plot = ancom_res,device = 'png',width = 6, height = 4)

```



# Correlation

## Formating
```{r}
# load data
immu_res <- read_csv(file = "ft_immu.csv")
immu_res[,c("trt", "immu", "pig", "time")] <- immu_res %>% select(trt, immu, pig, time) %>%  lapply(factor)
immu_res_nasal <- immu_res %>% filter(!is.na(time))
immu_res <- immu_res %>% filter(is.na(time)) %>% select(-time)

# combine metadata to immu data
metadata_immu <- metadata %>% filter(Experiment.id == 2) %>% filter(Pig.id %in% unique(immu_res$pig))
immu_full <- inner_join(metadata_immu, immu_res, by = c("Pig.id" = "pig")) 
immu_bal <- immu_full %>% filter(type == "BAL") %>% filter(immu %in% c("TBLN_CD3", "TBLN_CD3_172",  "BAL_CD3", "BAL_CD172", "TBLN_T_cells", "TBLN_T_helper",  "BAL_lgG_50", "BAL_lgG_100", "BAL_lgG_200",  "BAL_lgG_400", "BAL_lgG_800", "BAL_lgG_1600", "BAL_lgA_50", "BAL_lgA_100", "BAL_lgA_200", "BAL_lgA_400", "BAL_lgA_800", "BAL_lgA_1600"))
immu_ileum <- immu_full %>% filter(type == "Ileum")  %>% filter(!immu %in% c("TBLN_CD3", "TBLN_CD3_172",  "BAL_CD3", "BAL_CD172", "TBLN_T_cells", "TBLN_T_helper",  "BAL_lgG_50", "BAL_lgG_100", "BAL_lgG_200",  "BAL_lgG_400", "BAL_lgG_800", "BAL_lgG_1600", "BAL_lgA_50", "BAL_lgA_100", "BAL_lgA_200", "BAL_lgA_400", "BAL_lgA_800", "BAL_lgA_1600"))
immu_colon <- immu_full %>% filter(type == "Colon")  %>% filter(!immu %in% c("TBLN_CD3", "TBLN_CD3_172",  "BAL_CD3", "BAL_CD172", "TBLN_T_cells", "TBLN_T_helper",  "BAL_lgG_50", "BAL_lgG_100", "BAL_lgG_200",  "BAL_lgG_400", "BAL_lgG_800", "BAL_lgG_1600", "BAL_lgA_50", "BAL_lgA_100", "BAL_lgA_200", "BAL_lgA_400", "BAL_lgA_800", "BAL_lgA_1600"))
immu_fecal <- immu_full %>% filter(type == "Fecal")  %>% filter(!immu %in% c("TBLN_CD3", "TBLN_CD3_172",  "BAL_CD3", "BAL_CD172", "TBLN_T_cells", "TBLN_T_helper",  "BAL_lgG_50", "BAL_lgG_100", "BAL_lgG_200",  "BAL_lgG_400", "BAL_lgG_800", "BAL_lgG_1600", "BAL_lgA_50", "BAL_lgA_100", "BAL_lgA_200", "BAL_lgA_400", "BAL_lgA_800", "BAL_lgA_1600"))

# combine microbiota data to immu data for correlation
# BAL microbiota to TBLN,BAL & Nasal Swab
# fecal, colon, & Ileum to MLN
ps_rela <- transform_sample_counts(ps.rarefied, function(x) x / sum(x) ) 
ps_corr_fecal <- ps_rela %>% subset_samples(type == 'Fecal')
ps_corr_ileum <- ps_rela %>% subset_samples(type == 'Ileum')
ps_corr_colon <- ps_rela %>% subset_samples(type == 'Colon')
ps_corr_bal <- ps_rela %>% subset_samples(type == 'BAL')


# rarefied to genera level
ps_rela_corr_ileum_glom <- tax_glom(ps_corr_ileum, taxrank = 'Genus', NArm = FALSE)
ps_rela_corr_ileum_glom_melt <- psmelt(ps_rela_corr_ileum_glom)
ps_rela_corr_ileum_glom_melt$Genus <- as.character(ps_rela_corr_ileum_glom_melt$Genus)

ps_rela_corr_colon_glom <- tax_glom(ps_corr_colon, taxrank = 'Genus', NArm = FALSE)
ps_rela_corr_colon_glom_melt <- psmelt(ps_rela_corr_colon_glom)
ps_rela_corr_colon_glom_melt$Genus <- as.character(ps_rela_corr_colon_glom_melt$Genus)

ps_rela_corr_fecal_glom <- tax_glom(ps_corr_fecal, taxrank = 'Genus', NArm = FALSE)
ps_rela_corr_fecal_glom_melt <- psmelt(ps_rela_corr_fecal_glom)
ps_rela_corr_fecal_glom_melt$Genus <- as.character(ps_rela_corr_fecal_glom_melt$Genus)

ps_rela_corr_bal_glom <- tax_glom(ps_corr_bal, taxrank = 'Genus', NArm = FALSE)
ps_rela_corr_bal_glom_melt <- psmelt(ps_rela_corr_bal_glom)
ps_rela_corr_bal_glom_melt$Genus <- as.character(ps_rela_corr_bal_glom_melt$Genus)

# corr
corr_colon <- inner_join(immu_colon, ps_rela_corr_colon_glom_melt, by = c("sample.id", "Pig.id", "Experiment.id", "type", "Treatment", "time")) %>% 
  group_by(immu, Genus) %>% 
  summarise(p = cor.test(data,Abundance, method="kendall")$p.value, r = cor.test(data,Abundance, method="kendall")$estimate ) 
#NA appear if there are attributes with zero variance (with all elements equal), 0 in this case
corr_colon_p <- corr_colon %>% select(-r) %>%  pivot_wider(names_from = Genus, values_from = p)
corr_colon_p[is.na(corr_colon_p)] <- 1 
corr_colon_r <- corr_colon %>% select(-p) %>%  pivot_wider(names_from = Genus, values_from = r) 
corr_colon_r[is.na(corr_colon_r)] <- 0 


corr_ileum <- inner_join(immu_ileum, ps_rela_corr_ileum_glom_melt, by = c("sample.id", "Pig.id", "Experiment.id", "type", "Treatment", "time")) %>% 
  group_by(immu, Genus) %>% 
  summarise(p = cor.test(data,Abundance, method="kendall")$p.value, r = cor.test(data,Abundance, method="kendall")$estimate ) 
#NA appear if there are attributes with zero variance (with all elements equal), 0 in this case
corr_ileum_p <- corr_ileum %>% select(-r) %>%  pivot_wider(names_from = Genus, values_from = p)
corr_ileum_p[is.na(corr_ileum_p)] <- 1 
corr_ileum_r <- corr_ileum %>% select(-p) %>%  pivot_wider(names_from = Genus, values_from = r) 
corr_ileum_r[is.na(corr_ileum_r)] <- 0 


corr_fecal <- inner_join(immu_fecal, ps_rela_corr_fecal_glom_melt, by = c("sample.id", "Pig.id", "Experiment.id", "type", "Treatment", "time")) %>% 
  group_by(immu, Genus) %>% 
  summarise(p = cor.test(data,Abundance, method="kendall")$p.value, r = cor.test(data,Abundance, method="kendall")$estimate ) 
#NA appear if there are attributes with zero variance (with all elements equal), 0 in this case
corr_fecal_p <- corr_fecal %>% select(-r) %>%  pivot_wider(names_from = Genus, values_from = p)
corr_fecal_p[is.na(corr_fecal_p)] <- 1 
corr_fecal_r <- corr_fecal %>% select(-p) %>%  pivot_wider(names_from = Genus, values_from = r) 
corr_fecal_r[is.na(corr_fecal_r)] <- 0 


corr_bal <- inner_join(immu_bal, ps_rela_corr_bal_glom_melt, by = c("sample.id", "Pig.id", "Experiment.id", "type", "Treatment", "time")) %>% 
  group_by(immu, Genus) %>% 
  summarise(p = cor.test(data,Abundance, method="kendall")$p.value, r = cor.test(data,Abundance, method="kendall")$estimate ) 
#NA appear if there are attributes with zero variance (with all elements equal), 0 in this case
corr_bal_p <- corr_bal %>% select(-r) %>%  pivot_wider(names_from = Genus, values_from = p)
corr_bal_p[is.na(corr_bal_p)] <- 1 
corr_bal_r <- corr_bal %>% select(-p) %>%  pivot_wider(names_from = Genus, values_from = r) 
corr_bal_r[is.na(corr_bal_r)] <- 0 

corr_bal_p

```

## Get significant taxa (if in any of the immunology data)
```{r}
sig_taxa <- function(x){
 taxa <- c()
  for (i in 1:ncol(x)){
    if (any(x[i] < 0.05)){
      taxa <- append(names(x)[i], taxa)
    }
  }
    return(taxa)
}

sig_taxa_bal <- corr_bal_p %>% ungroup() %>% select(-immu) %>% sig_taxa()
sig_taxa_colon <- corr_colon_p %>% ungroup() %>% select(-immu) %>% sig_taxa()
sig_taxa_ileum <- corr_ileum_p %>% ungroup() %>% select(-immu) %>% sig_taxa()
sig_taxa_fecal <- corr_fecal_p %>% ungroup() %>% select(-immu) %>% sig_taxa()
```


## Plot correlation results
### Formating
```{r}
corr_colon_r_matrix_pre <- corr_colon_r[c("immu", sig_taxa_colon)]
corr_colon_r_matrix_pre[corr_colon_p[c("immu", sig_taxa_colon)] > 0.05] <- 0 # if not significant (p > 0.05), set r to 0

corr_fecal_r_matrix_pre <- corr_fecal_r[c("immu", sig_taxa_fecal)]
corr_fecal_r_matrix_pre[corr_fecal_p[c("immu", sig_taxa_fecal)] > 0.05] <- 0 # if not significant (p > 0.05), set r to 0

corr_bal_r_matrix_pre <- corr_bal_r[c("immu", sig_taxa_bal)]
corr_bal_r_matrix_pre[corr_bal_p[c("immu", sig_taxa_bal)] > 0.05] <- 0 # if not significant (p > 0.05), set r to 0

# combine colon and fecal matrix, if not significant set r to 0
corr_colon_r_matrix_pre[setdiff(colnames(corr_fecal_r_matrix_pre), colnames(corr_colon_r_matrix_pre))] <- 0
corr_fecal_r_matrix_pre[setdiff(colnames(corr_colon_r_matrix_pre), colnames(corr_fecal_r_matrix_pre))] <- 0
corr_colon_r_matrix_pre["microbiota"] <- "colon"
corr_fecal_r_matrix_pre["microbiota"] <- "fecal"

corr_dig_column_name <- union(colnames(corr_fecal_r_matrix_pre), colnames(corr_colon_r_matrix_pre))
corr_dig_for_heatmap <- rbind(corr_colon_r_matrix_pre[corr_dig_column_name], corr_fecal_r_matrix_pre[corr_dig_column_name]) %>% ungroup()
corr_dig_for_heatmap <- corr_dig_for_heatmap %>% 
  plyr::rename(c("Pseudoramibacter_Eubacterium" = "Pseudoramibacter", "Unclassified [Mogibacteriaceae]" = "Unclassified Mogibacteriaceae", "Unclassified [Barnesiellaceae]" = "Unclassified Barnesiellaceae", "[Ruminococcus]" = "Ruminococcus", "[Eubacterium]" = "Eubacterium")) # formating; plyr::rename functions as pandas 


```

### Complexheatmap, fecal and colon
```{r}

corr_dig_for_heatmap$immu <- str_split_fixed(corr_dig_for_heatmap$immu, "MLN_", 2)[,2]
replacement <- c("CD172" = "Myeloid Cells", "CD3$" = "Non-T cells", "T_helper" = "T-helper/memory cells", "T_cells" = "Cytotoxic T cells")
corr_dig_for_heatmap$immu <- str_replace_all(corr_dig_for_heatmap$immu, replacement)
# visualizing with complexheatmap package
row_ha <- rowAnnotation( Immunology = corr_dig_for_heatmap$immu,
                         col = list(Immunology = c("Myeloid Cells" = "#ebc034", "Non-T cells" = "#34eb58", "T-helper/memory cells" = "#080808", "Cytotoxic T cells" = "#c73c9d"))
                         ) 

heatmap_dig <- corr_dig_for_heatmap %>% select(-c(immu, microbiota)) %>% Heatmap(name = "r", 
                                                                  #left_annotation = row_ha, # annotation  
                                                                  row_split = corr_dig_for_heatmap$microbiota, # split colon and fecal 
                                                                  row_labels = corr_dig_for_heatmap$immu,
                                                                  show_heatmap_legend = FALSE)

# draw(heatmap_dig, heatmap_legend_side = "bottom", annotation_legend_side = "bottom") # adjust legend position
heatmap_dig

```


### Complexheatmap, Bal 
```{r}
corr_bal_r_matrix_pre <- corr_bal_r[c("immu", sig_taxa_bal)]
corr_bal_r_matrix_pre[corr_bal_p[c("immu", sig_taxa_bal)] > 0.05] <- 0 # if not significant (p > 0.05), set r to 0
corr_bal_r_matrix_pre <- corr_bal_r_matrix_pre %>% plyr::rename(c("[Ruminococcus]" =  "Ruminococcus"))

# rename, for better visualization 
corr_bal_r_matrix_row_rename <- c("Myeloid cells", "Non-T cells",
                                  "1 in 100", "1 in 1600", "1 in 200", "1 in 400", "1 in 50", "1 in 800",
                                   "1 in 100", "1 in 1600", "1 in 200", "1 in 400", "1 in 50", "1 in 800",
                                  "Non-T cells", "Myeloid cells",
                                  "T cells", "T helper/memory cells")

corr_bal_r_matrix_pre$immu <- corr_bal_r_matrix_row_rename

corr_bal_r_matrix_pre$Cells <- c("BAL Cells Myloid Cells", "BAL Cells Myloid Cells", rep("lgA", 6), rep("lgG", 6), "TBLN MNCs Myloid cells", "TBLN MNCs Myloid cells", "TBLN MNCs Immune Cells", "TBLN MNCs Immune Cells")

corr_bal_r_matrix_pre <- corr_bal_r_matrix_pre %>% mutate(immu = as.character(immu)) %>%  arrange(Cells, nchar(immu))  # nchar()  
corr_bal_r_matrix_pre$immu <- corr_bal_r_matrix_pre$immu %>% factor(levels = c("1 in 50", "1 in 100", "1 in 200", "1 in 400", "1 in 800", "1 in 1600", "Non-T cells", "Myeloid cells", "T cells", "T helper/memory cells"))
corr_bal_r_matrix_pre$order = 1:nrow(corr_bal_r_matrix_pre)


row_ha_bal <- rowAnnotation(Cells = corr_bal_r_matrix_pre$Cells, 
                             col = list(Cells = c("BAL Cells Myloid Cells" = "#ebc034", "lgA" = "#34eb58", "lgG" = "#080808", "TBLN MNCs Immune Cells" = "#c73c9d", "TBLN MNCs Myloid cells" = "#3c94c7")))                          
heatmap_bal <- corr_bal_r_matrix_pre %>% ungroup() %>% select(-c(immu,  Cells, order)) %>% Heatmap(name = "r", row_split = corr_bal_r_matrix_pre$Cells, row_title = NULL, left_annotation = row_ha_bal, row_order = corr_bal_r_matrix_pre$order, row_labels = corr_bal_r_matrix_pre$immu)

heatmap_bal_adjusted <- draw(heatmap_bal, heatmap_legend_side = "bottom", annotation_legend_side = "bottom", newpage = FALSE)
```

### Combine correlation results as a single figure
```{r}
# use `grid.grabExpr()` to capture the output of ComplexHeatmap as a `grob` object and arrange `grob` objects by other package.
# see https://support.bioconductor.org/p/103113/
grob1 <-grid.grabExpr(draw(heatmap_dig, ))
grob2 <-grid.grabExpr(draw(heatmap_bal_adjusted))

heatmap_dig_ready <- annotate_figure(grob1, fig.lab = "(a)")
heatmap_bal_ready <- annotate_figure(grob2, fig.lab = "(b)")
corr_heatmap <- ggarrange(heatmap_dig_ready, heatmap_bal_ready, nrow = 2, heights = c(4,6))



ggsave('res_ex2/correlation.png', plot = corr_heatmap,device = 'png',width = 7, height = 12)
```
